Automated Residual Plot Assessment with the R Package autovi and Shiny App autovi.web

### 2024-11-18 autovi_paper.tex
drastic paper trimming

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

@@

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{anzsj}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Software for Automated Residual Plot Assessment: autovi and autovi.web},
  pdfauthor={Weihao Li; Dianne Cook; Emi Tanaka; Susan VanderPlas; Klaus Ackermann},
  pdfkeywords={initial data analysis, statistical graphics, data
visualization, visual inference, computer vision, machine
learning, hypothesis testing, regression analysis},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\runningheads{
Software for Automated Residual Plot Assessment
}{
\firstletters{WEIHAO}LI, \firstletters{DIANNE}COOK, \firstletters{EMI}TANAKA, \firstletters{SUSAN}VANDERPLAS, AND  \firstletters{KLAUS}ACKERMANN
}

\title{Software for Automated Residual Plot Assessment: autovi and
autovi.web}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{ANZJS Quarto Template}

@@

The main purpose of \texttt{autovi} is to provide rejection decisions
and \(p\)-values for testing whether a regression model is correctly
specified. The package introduces a novel approach to automating
statistical analysis, particularly in the interpretation of residual
plots. The name \texttt{autovi} stands for automated visual inference.

@@

\textbf{?@fig-three-examples} shows three sets of plots of residuals
against fitted values. The simulated example in (a) might be interpreted
as a heteroscedastic pattern, however the automated reading would
predict this to have a visual signal strength (VSS) of 1.53, with a
corresponding \(p\)-value of 0.25. This means it is would be interpreted
as a good residual plot, that there is nothing in the data to indicate a
violation of model assumptions. Skewness in the predictor variables is
generating the apparent heteroscedasticity, where the smaller variance
in residuals at larger fitted values is due to smaller sample size only.
The Breusch-Pagan test \citep{breusch1979simple} for heteroscedasticity
would also not reject this as good residual plot.

@@

{\centering \includegraphics[width=1\textwidth,height=\textheight]{autovi_paper_files/figure-pdf/three-examples-1.pdf}

@@

\caption{Reading residual plots can be a difficult task, particular for
students new to statistical modeling. The autovi package makes it
easier. Here are three examples of residual plots, which may appear to
have structure. According to autovi, the visual signal strengths (VSS)
of these three examples are approximately (a) 1.53, (b) 3.57, (c) 5.87,
resulting in (b), (c) being significant violations of good residuals,
but (a) is consistent with a good residual plot.}

@@

The \texttt{autovi} package is available on CRAN. It is actively
developed and maintained, with the latest updates accessible on GitHub
at \url{https://github.com/TengMCing/autovi}. The code discussed in this
paper is based on \texttt{autovi} version 0.4.1.

@@

\subsection{Usage}\label{sec-autovi-usage}

To get started quickly, only three function calls are needed to obtain a
summary of the automated residual assessment:

@@

The three functions are explained as follows:

The summary reports key findings such as the VSS of the true residual
plot and the \(p\)-value of the automated visual test. The \(p\)-value
is the ratio of null plots that have VSS greater than or equal to that
of the true residual plot. We typically reject the null hypothesis when
the \(p\)-value is smaller than or equal to a desired significance
level, such as 5\%. The report also provides sample quantiles of VSS for
null and bootstrapped plots, helping to explain the severity and
likelihood of model violations.

Although the \(p\)-value is sufficient for automated decision-making,
users can visually inspect the original residual plot alongside a sample
null plot. This visual comparison can clarify why \(H_0\) is either
rejected or not, and help identify potential remedies. The
\texttt{plot\_pair()} and \texttt{plot\_lineup()} methods facilitate
this comparison.

@@

\includegraphics[width=1\textwidth,height=\textheight]{autovi_paper_files/figure-pdf/fig-plot-pair-1.pdf}

@@

\includegraphics[width=1\textwidth,height=\textheight]{autovi_paper_files/figure-pdf/fig-summary-plot-1.pdf}

@@

\includegraphics[width=1\textwidth,height=\textheight]{autovi_paper_files/figure-pdf/fig-autovi-diag-1.png}

@@

The modules for VSS prediction and \(p\)-value computation are
predefined and cannot be overridden, although users can interact with
them directly through function arguments. Similarly, the image reading
and resizing module is fixed but will adapt to different Keras models by
checking their input shapes. The remaining seven modules are designed to
be overridable, enabling users to tailor the infrastructure to their
specific needs. These modules are discussed in detail in the following
sections.

\subsubsection{Initialization}\label{initialization}

An \texttt{autovi} checker can be initialized by supplying two primary
inputs, including a regression model object, such as an \texttt{lm}
object representing the result of a linear regression model, and a
trained computer vision model compatible with the \texttt{Keras}
\citep{chollet2015keras} Application Programming Interface (API), to the
\texttt{AUTO\_VI} class constructor \texttt{auto\_vi()}. The
\texttt{residual\_checker()} introduced in
Section~\ref{sec-autovi-usage} is a thin wrapper around
\texttt{auto\_vi()}, which will call \texttt{get\_keras\_model()} during
initialization. \texttt{get\_keras\_model()} is a function to download a
trained computer vision model (described in \citet{li2024automated})
from GitHub. ``vss\_phn\_32'' specifies a model that predicts VSS and is
trained on residuals with polynomial, heteroscedasticity, and
non-normality patterns (phn). More details about the hosted models will
be provided in Section~\ref{sec-trained-model-hosting}.

The input of the constructor will be stored as attributes of the checker
and can be accessed by the user through the \texttt{\$} operator.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(autovi)}
\NormalTok{checker }\OtherTok{\textless{}{-}} \FunctionTok{auto\_vi}\NormalTok{(}\AttributeTok{fitted\_model =} \FunctionTok{lm}\NormalTok{(dist }\SpecialCharTok{\textasciitilde{}}\NormalTok{ speed, }\AttributeTok{data =}\NormalTok{ cars), }
                   \AttributeTok{keras\_model =} \FunctionTok{get\_keras\_model}\NormalTok{(}\StringTok{"vss\_phn\_32"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

Optionally, the user may specify the node index of the output layer of
the trained computer vision model to be monitored by the checker via the
\texttt{node\_index} argument if there are multiple output nodes. This
is particularly useful for multiclass classifiers when the user wants to
use one of the nodes as a VSS indicator.

After initializing the object, you can print the checker to view its
status.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{── }\SpecialCharTok{\textless{}}\NormalTok{AUTO\_VI object}\SpecialCharTok{\textgreater{}}
\NormalTok{Status}\SpecialCharTok{:}
 \SpecialCharTok{{-}}\NormalTok{ Fitted model}\SpecialCharTok{:}\NormalTok{ lm}
 \SpecialCharTok{{-}}\NormalTok{ Keras model}\SpecialCharTok{:}\NormalTok{ (None, }\DecValTok{32}\NormalTok{, }\DecValTok{32}\NormalTok{, }\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}\NormalTok{ (None, }\DecValTok{5}\NormalTok{) }\OtherTok{{-}\textgreater{}}\NormalTok{ (None, }\DecValTok{1}\NormalTok{)}
    \SpecialCharTok{{-}}\NormalTok{ Output node index}\SpecialCharTok{:} \DecValTok{1}
 \SpecialCharTok{{-}}\NormalTok{ Result}\SpecialCharTok{:}\NormalTok{ UNKNOWN }
\end{Highlighting}
\end{Shaded}

The status includes the list of regression model classes (as provided by
the built-in \texttt{class()} function), the input and output shapes of
the Keras model in the standard \texttt{Numpy} format
\citep{harris2020array}, the output node index being monitored, and the
assessment result. If no check has been run yet, the assessment result
will display as ``UNKNOWN''.

\subsubsection{Fitted Values and Residuals
Extraction}\label{fitted-values-and-residuals-extraction}

To be able to predict VSS for a residual plot, both fitted values and
residuals are needed to be extracted from the regression model object
supplied by the user. In R, statistical models like \texttt{lm} (linear
model) and \texttt{glm} (generalized linear model) typically support the
use of generic functions such as \texttt{fitted()} and \texttt{resid()}
to retrieve these values. The \texttt{get\_fitted\_and\_resid()} method,
called by the checker, relies on these generic functions by default.
However, generic functions only work with classes that have appropriate
method implementations. Some regression modeling packages may not fully
adhere to the \texttt{stats} package guidelines for implementing these
functions. In such cases, overriding the method becomes necessary.

By design, the \texttt{get\_fitted\_and\_resid()} method accepts a
regression model object as input and returns a \texttt{tibble} (a modern
presentation of the \texttt{data.frame}) with two columns:
\texttt{.fitted} and \texttt{.resid}, representing the fitted values and
residuals, respectively. If no input is supplied, the method uses the
regression model object stored in the checker. Although modules in the
\texttt{autovi} infrastructure make minimal assumptions about other
modules, they do require strictly defined input and output formats to
ensure data validation and prevent fatal bugs. Therefore, any overridden
method should follow to these conventions.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{checker}\SpecialCharTok{$}\FunctionTok{get\_fitted\_and\_resid}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 50 x 2
   .fitted .resid
     <dbl>  <dbl>
 1   -1.85   3.85
 2   -1.85  11.8 
 3    9.95  -5.95
 4    9.95  12.1 
 5   13.9    2.12
 6   17.8   -7.81
 7   21.7   -3.74
 8   21.7    4.26
 9   21.7   12.3 
10   25.7   -8.68
# i 40 more rows
\end{verbatim}

\subsubsection{Data Extraction}\label{data-extraction}

For linear regression model in R, the model frame contains all the data
required by a formula for evaluation. This is essential for
bootstrapping and refitting the model when constructing a bootstrapped
distribution of VSS. Typically, the model frame can be extracted from
the regression model object using the \texttt{model.frame()} generic
function, which is the default method used by \texttt{get\_data()}.
However, some regression models do not use a formula or are evaluated
differently, potentially lacking a model frame. In such cases, users can
either provide the data used to fit the regression model through the
\texttt{data} argument when constructing the checker, or customize the
method to better suit their needs. It's worth noting that this module is
only necessary if bootstrapping is required, as the model frame is not
used in other modules of the infrastructure.

The \texttt{get\_data()} method accepts a regression model object as
input and returns a \texttt{data.frame} representing the model frame of
the fitted regression model. If no input is supplied, the regression
model stored in the checker will be used.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{checker}\SpecialCharTok{$}\FunctionTok{get\_data}\NormalTok{() }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{head}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
  dist speed
1    2     4
2   10     4
3    4     7
4   22     7
5   16     8
6   10     9
\end{verbatim}

\subsubsection{Bootstrapping and Model
Refitting}\label{bootstrapping-and-model-refitting}

Bootstrapping a regression model typically involves sampling the
observations with replacement and refitting the model with the
bootstrapped data. The \texttt{boot\_method()} method follows this
bootstrapping scheme by default. It accepts a fitted regression model
and a \texttt{data.frame} as inputs, and returns a \texttt{tibble} of
bootstrapped residuals. If no inputs are provided, the method uses the
regression model stored in the checker and the result of the
\texttt{get\_data()} method.

Note that instead of calling \texttt{get\_data()} implicitly within the
method, it is used as part of the default argument definition. This
approach allows users to bypass the \texttt{get\_data()} method entirely
and directly supply a \texttt{data.frame} to initiate the bootstrap
process. Many other methods in \texttt{autovi} adopt this principle when
possible, where dependencies are explicitly listed in the formal
arguments. This design choice enhances the reusability and isolation of
modules, offers better control for testing, and simplifies the overall
process.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{checker}\SpecialCharTok{$}\FunctionTok{boot\_method}\NormalTok{(}\AttributeTok{data =}\NormalTok{ checker}\SpecialCharTok{$}\FunctionTok{get\_data}\NormalTok{())}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 50 x 2
   .fitted .resid
     <dbl>  <dbl>
 1   27.0   -2.96
 2   38.8  -12.8 
 3   34.8   -8.82
 4   27.0  -13.0 
 5   11.2    4.76
 6   42.7   -2.68
 7   42.7   -2.68
 8   38.8  -18.8 
 9   38.8   15.2 
10   -4.47   6.47
# i 40 more rows
\end{verbatim}

\subsubsection{Auxiliary Computation}\label{auxiliary-computation}

As described in \citet{li2024automated}, in some cases, a residual plot
alone may not provide enough information to accurately determine VSS.
For instance, when the points in the residual plot have significant
overlap, the trend and shape of the residual pattern can be difficult to
discern. Including auxiliary variables, such as the number of
observations, as additional inputs to the computer vision model can be
beneficial. To address this, \texttt{autovi} includes internal functions
within the checker that automatically detect the number of inputs
required by the provided Keras model. If multiple inputs are necessary,
the checker invokes the \texttt{auxiliary()} method to compute these
additional inputs.

The \texttt{auxiliary()} method takes a \texttt{data.frame} containing
fitted values and residuals as input and returns a \texttt{data.frame}
with five numeric columns. These columns represent four scagnostics ---
``Monotonic'', ``Sparse'', ``Striped'', and ``Splines'' --- calculated
using the \texttt{cassowaryr} package, as well as the number of
observations. This approach is consistent with the training process of
the computer vision models described in \citet{li2024automated}. If no
\texttt{data.frame} is provided, the method will default to retrieving
fitted values and residuals by calling
\texttt{get\_fitted\_and\_resid()}.

Technically, any Keras-implemented computer vision model can be adapted
to accept an image as the primary input and additional variables as
secondary inputs by adding a data pre-processing layer before the actual
input layer. If users wish to override \texttt{auxiliary()}, the output
should be a \texttt{data.frame} with a single row and the number of
columns such that its concatenation matches the number of parameters for
the corresponding layer in the supplied Keras model.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{checker}\SpecialCharTok{$}\FunctionTok{auxiliary}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 1 x 5
  measure_monotonic measure_sparse measure_splines measure_striped     n
              <dbl>          <dbl>           <dbl>           <dbl> <int>
1            0.0621          0.470          0.0901            0.62    50
\end{verbatim}

\subsubsection{Null Residual Simulation}\label{sec-autovi-null-method}

A fundamental element of the automated residual assessment described in
\citet{li2024automated} is comparing the VSS of null plots with that of
the true residual plot. However, due to the variety of regression
models, there is no universal method for simulating null residuals that
are consistent with model assumptions. Fortunately, for classical normal
linear regression models, null residuals can be effectively simulated
using the residual rotation method, as outlined in
\citet{buja2009statistical}. This process involves generating random
draws from a standard normal distribution, regressing these draws on the
original predictors, and then rescaling the resulting residuals by the
ratio of the residual sum of squares to the that of the original linear
regression model. Other regression models, such as \texttt{glm}
(generalized linear model) and \texttt{gam} (generalized additive
model), generally cannot use this method to efficiently simulate null
residuals. Therefore, it is recommended that users override the
\texttt{null\_method()} to suit their specific model. The
\texttt{null\_method()} takes a fitted regression model as input,
defaulting to the regression model stored in the checker, and returns a
\texttt{tibble}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{checker}\SpecialCharTok{$}\FunctionTok{null\_method}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 50 x 2
   .fitted  .resid
     <dbl>   <dbl>
 1   -1.85  18.2  
 2   -1.85  -0.765
 3    9.95 -12.8  
 4    9.95  18.6  
 5   13.9    2.57 
 6   17.8    7.03 
 7   21.7  -11.1  
 8   21.7  -13.2  
 9   21.7  -12.6  
10   25.7    3.57 
# i 40 more rows
\end{verbatim}

\subsubsection{Plotting}\label{plotting}

Plotting is a crucial aspect of residual plot diagnostics because
aesthetic elements like marker size, marker color, and auxiliary lines
impact the presentation of information. There are computer vision models
trained to handle images captured in various scenarios. For example, the
VGG16 model \citep{simonyan2014very} can classify objects in images
taken under different lighting conditions and is robust to image
rotation. However, data plots are a special type of image as the
plotting style can always be consistent if controlled properly.
Therefore, we assume computer vision models built for reading residual
plots will be trained with residual plots of a specific aesthetic style.
In this case, it is best to predict plots using the same style for
optimal performance. The plotting method \texttt{plot\_resid()} handles
this aspect.

\texttt{plot\_resid()} accepts a \texttt{data.frame} containing fitted
values and residuals, along with several customization options: a
\texttt{ggplot} theme, an \texttt{alpha} value to control the
transparency of data points, a \texttt{size} value to set the size of
data points, and a \texttt{stroke} value to define the thickness of data
point edges. Additionally, it includes four Boolean arguments to toggle
the display of axes, legends, grid lines, and a horizontal red line. By
default, it replicates the style we used to generate the training
samples for the computer vision models described in
\citet{li2024automated}. In brief, the residual plot omits axis text and
ticks, titles, and background grid lines, featuring only a red line at
\(y = 0\). It retains only the necessary components of a residual plot.
If the computer vision model is trained with a different but consistent
aesthetic style, \texttt{plot\_resid()} should be overridden.

The method returns a \texttt{ggplot} object
(Figure~\ref{fig-autovi-plot-resid}), which can be saved as a PNG file
in the following module. If no data is provided, the method will use
\texttt{get\_fitted\_and\_resid()} to retrieve the fitted values and
residuals from the regression model stored in the checker.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{checker}\SpecialCharTok{$}\FunctionTok{plot\_resid}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics[width=1\textwidth,height=\textheight]{autovi_paper_files/figure-pdf/fig-autovi-plot-resid-1.pdf}

\caption{\label{fig-autovi-plot-resid}Residual plot of the regression
model stored in the checker.}

To manually generate true residual plots, null plots, or bootstrapped
residual plots, you can pass the corresponding \texttt{data.frame}
produced by the \texttt{get\_fitted\_and\_resid()},
\texttt{null\_method()}, and \texttt{boot\_method()} methods to the
\texttt{plot\_resid()} method, respectively.

\subsubsection{Plot Saving}\label{plot-saving}

Another key aspect of a standardized residual plot is its resolution. In
\citet{li2024automated}, we used an image format of 420 pixels in height
and 525 pixels in width. This resolution was chosen because the original
set, consisting of 20 residual plots arranged in a four by five grid,
was represented by an image of 2100 by 2100 pixels. The
\texttt{save\_plot()} method accepts a \texttt{ggplot} object as input
and saves it as a PNG file to the location specified by the
\texttt{path} argument. If no path is provided, the PNG file is saved to
a temporary file.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{checker}\SpecialCharTok{$}\FunctionTok{plot\_resid}\NormalTok{() }\SpecialCharTok{|\textgreater{}} 
\NormalTok{  checker}\SpecialCharTok{$}\FunctionTok{save\_plot}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] "/var/folders/61/bv7_1qzs20x6fjb2rsv7513r0000gn/T//Rtmp6Lce76/file135f86b4a7b69.png"
\end{verbatim}

\subsubsection{Image Reading and
Resizing}\label{image-reading-and-resizing}

When training computer vision models, it is common to test various input
sizes for the same architecture to identify the optimal setup. This
involves preparing the original training image at a higher resolution
than required and then resizing it to match the input size during
training. The \texttt{autovi} package includes a class,
\texttt{KERAS\_WRAPPER}, to simplify this process. This Keras wrapper
class features a method called \texttt{image\_to\_array()}, which reads
an image as a \texttt{PIL} image using the \texttt{pillow} Python
package, resizes it to the target input size required by the Keras
model, and converts it to a \texttt{Numpy} array.

To construct a \texttt{KERAS\_WRAPPER} object, you need to provide the
Keras model as the main argument. However, users generally do not need
to interact with this class directly, as the \texttt{autovi} checker
automatically invokes its methods when performing VSS predictions. The
\texttt{image\_to\_array()} method takes the path to the image file, the
target height, and the target width as inputs and returns a
\texttt{Numpy} array. If not specified, the target height and target
width will be retrieved from the input layer of the Keras model by the
\texttt{get\_input\_height()} and \texttt{get\_input\_width()} method of
\texttt{KERAS\_WRAPPER}.

The following code example demonstrate the way to manually generate the
true residual plot, save it as PNG file, and load it back as
\texttt{Numpy} array.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{wrapper }\OtherTok{\textless{}{-}} \FunctionTok{keras\_wrapper}\NormalTok{(}\AttributeTok{keras\_model =}\NormalTok{ checker}\SpecialCharTok{$}\NormalTok{keras\_model)  }
\NormalTok{input\_array }\OtherTok{\textless{}{-}}\NormalTok{ checker}\SpecialCharTok{$}\FunctionTok{plot\_resid}\NormalTok{() }\SpecialCharTok{|\textgreater{}} 
\NormalTok{  checker}\SpecialCharTok{$}\FunctionTok{save\_plot}\NormalTok{() }\SpecialCharTok{|\textgreater{}}
\NormalTok{  wrapper}\SpecialCharTok{$}\FunctionTok{image\_to\_array}\NormalTok{()}
\NormalTok{input\_array}\SpecialCharTok{$}\NormalTok{shape}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
(1, 32, 32, 3)
\end{verbatim}

\subsubsection{Visual Signal Strength (VSS)
Prediction}\label{visual-signal-strength-vss-prediction}

VSS, as discussed in \citet{li2024automated}, estimates the distance
between the input residual plot and a theoretically good residual plot.
It can be defined in various ways, much like different methods for
measuring the distance between two points. This will not impact the
\texttt{autovi} infrastructure as long as the provided Keras model can
predict the intended measure.

There are several ways to obtain VSS from the checker, with the most
direct being the \texttt{vss()} method. By default, this method predicts
the VSS for the true residual plot. If a \texttt{ggplot} or a
\texttt{data.frame}, such as null residuals generated by the
\texttt{null\_method()}, is explicitly provided, the method will use
that input to predict VSS accordingly. Note that if a \texttt{ggplot} is
provided, auxiliary inputs must be supplied manually via the
\texttt{auxiliary} argument, as we assume that auxiliary variables can
not be computed directly from a \texttt{ggplot}.

Another way to obtain VSS is by calling the \texttt{check()} method.
This comprehensive method perform extensive diagnostics on the true
residual plot and store the VSS in the \texttt{check\_result} field of
the checker. Additionally, for obtaining VSS for null residual plots and
bootstrapped residual plots, there are two specialized methods,
\texttt{null\_vss()} and \texttt{boot\_vss()}, designed for this purpose
respectively.

Calling the \texttt{vss()} method without arguments will predict the VSS
for the true residual plot and return the result as a single-element
\texttt{tibble}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{checker}\SpecialCharTok{$}\FunctionTok{vss}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 1 x 1
    vss
  <dbl>
1  3.16
\end{verbatim}

Providing a \texttt{data.frame} of null residuals or a null residual
plot yields the same VSS.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{null\_resid }\OtherTok{\textless{}{-}}\NormalTok{ checker}\SpecialCharTok{$}\FunctionTok{null\_method}\NormalTok{()}
\NormalTok{checker}\SpecialCharTok{$}\FunctionTok{vss}\NormalTok{(null\_resid)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 1 x 1
    vss
  <dbl>
1  1.23
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{null\_resid }\SpecialCharTok{|\textgreater{}}
\NormalTok{  checker}\SpecialCharTok{$}\FunctionTok{plot\_resid}\NormalTok{() }\SpecialCharTok{|\textgreater{}}
\NormalTok{  checker}\SpecialCharTok{$}\FunctionTok{vss}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 1 x 1
    vss
  <dbl>
1  1.23
\end{verbatim}

The \texttt{null\_vss()} helper method primarily takes the number of
null plots as input. If the user wants to use a ad hoc null simulation
scheme, it can be provided via the \texttt{null\_method} argument.
Intermediate results, including null residuals and null plots, can be
returned by enabling \texttt{keep\_null\_data} and
\texttt{keep\_null\_plot}. The VSS, along with null residuals and null
plots, will be stored in a \texttt{tibble} with three columns. The
following code example demonstrates how to predict the VSS for five null
residual plots while keeping the intermediate results.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{checker}\SpecialCharTok{$}\FunctionTok{null\_vss}\NormalTok{(}\DecValTok{5}\NormalTok{L, }
                 \AttributeTok{keep\_null\_data =} \ConstantTok{TRUE}\NormalTok{, }
                 \AttributeTok{keep\_null\_plot =} \ConstantTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 5 x 3
    vss data              plot  
  <dbl> <list>            <list>
1 0.982 <tibble [50 x 2]> <gg>  
2 1.58  <tibble [50 x 2]> <gg>  
3 2.09  <tibble [50 x 2]> <gg>  
4 0.742 <tibble [50 x 2]> <gg>  
5 2.21  <tibble [50 x 2]> <gg>  
\end{verbatim}

The \texttt{boot\_vss()} helper method is similar to
\texttt{null\_vss()}, with some differences in argument names. The
following code example demonstrates how to predict the VSS for five
bootstrapped residual plots while keeping the intermediate results.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{checker}\SpecialCharTok{$}\FunctionTok{boot\_vss}\NormalTok{(}\DecValTok{5}\NormalTok{L,}
                 \AttributeTok{keep\_boot\_data =} \ConstantTok{TRUE}\NormalTok{,}
                 \AttributeTok{keep\_boot\_plot =} \ConstantTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 5 x 3
    vss data              plot  
  <dbl> <list>            <list>
1  3.12 <tibble [50 x 2]> <gg>  
2  2.37 <tibble [50 x 2]> <gg>  
3  2.87 <tibble [50 x 2]> <gg>  
4  1.85 <tibble [50 x 2]> <gg>  
5  3.66 <tibble [50 x 2]> <gg>  
\end{verbatim}

\subsubsection{\texorpdfstring{\(p\)-value
Computation}{p-value Computation}}\label{p-value-computation}

>>>>

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

@@

\usepackage[]{natbib}
\bibliographystyle{anzsj}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Automated Residual Plot Assessment with the R Package autovi and Shiny App autovi.web},
  pdfauthor={Weihao Li; Dianne Cook; Emi Tanaka; Susan VanderPlas; Klaus Ackermann},
  pdfkeywords={initial data analysis, statistical graphics, data
visualization, visual inference, computer vision, machine
learning, hypothesis testing, regression analysis},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\runningheads{
Automated Residual Plot Assessment
}{
\firstletters{WEIHAO}LI, \firstletters{DIANNE}COOK, \firstletters{EMI}TANAKA, \firstletters{SUSAN}VANDERPLAS, AND  \firstletters{KLAUS}ACKERMANN
}

\title{Automated Residual Plot Assessment with the R Package autovi and
Shiny App autovi.web}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{ANZJS Quarto Template}

@@

The main purpose of \texttt{autovi} is to provide rejection decisions
and \(p\)-values for testing the null hypothesis (\(H_0\)) that the
regression model is correctly specified. The package provides automated
interpretation of residual plots using computer vision. The name
\texttt{autovi} stands for \textbf{auto}mated \textbf{v}isual
\textbf{i}nference.

@@

Figure~\ref{fig-three-examples} shows three sets of plots of residuals
against fitted values. The simulated example in (a) might be interpreted
as a heteroscedastic pattern, however the automated reading would
predict this to have a visual signal strength (VSS) of 1.53, with a
corresponding \(p\)-value of 0.25. This means it is would be interpreted
as a good residual plot, that there is nothing in the data to indicate a
violation of model assumptions. Skewness in the predictor variables is
generating the apparent heteroscedasticity, where the smaller variance
in residuals at larger fitted values is due to smaller sample size only.
The Breusch-Pagan test \citep{breusch1979simple} for heteroscedasticity
would also not reject this as good residual plot.

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{autovi_paper_files/figure-pdf/fig-three-examples-1.pdf}

@@

\caption{\label{fig-three-examples}Reading residual plots can be a
difficult task, particular for students new to statistical modeling. The
autovi package makes it easier. Here are three examples of residual
plots, which may appear to have structure. According to autovi, the
visual signal strengths (VSS) of these three examples are approximately
(a) 1.53, (b) 3.57, (c) 5.87, resulting in (b), (c) being significant
violations of good residuals, but (a) is consistent with a good residual
plot.}

@@

The \texttt{autovi} package is available on CRAN. It is actively
developed and maintained, with the latest updates accessible on GitHub.
The code discussed in this paper is based on \texttt{autovi} version
0.4.1.

@@

We recommend using the Shiny app \texttt{autovi.web} if encountering
installation problems.

\subsection{Numerical Summary}\label{sec-autovi-usage}

Three steps are needed to get an automated assessment of a set of
residuals and fitted values:

The code to do this is:

@@

The summary includes observed VSS of the true residual plot and
associated \(p\)-value of the automated visual test. The \(p\)-value is
the proportion of null plots (out of the total 100) that have VSS
greater than or equal to that of the true residual plot. The report also
provides sample quantiles of VSS for null samples and bootstrapped data
plots, providing more information about the sampling variability and a
likelihood of model violations. The likelihood is computed from the
proportion of values greater than the observed VSS in both the
bootstrapped data values and the simulated null values.

\subsection{Visual Summary}\label{sec-autovi-visual}

Users can visually inspect the original residual plot alongside a sample
null plot using \texttt{plot\_pair()} or a lineup of null plot
\texttt{plot\_lineup()}. This visual comparison can clarify why \(H_0\)
is either rejected or not, and help identify potential remedies.

@@

\includegraphics[width=0.6\linewidth,height=\textheight,keepaspectratio]{autovi_paper_files/figure-pdf/fig-plot-pair-1.pdf}

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{autovi_paper_files/figure-pdf/fig-summary-plot-1.pdf}

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{autovi_paper_files/figure-pdf/fig-autovi-diag-1.png}

@@

The modules for VSS prediction and \(p\)-value computation are
predefined and cannot be overridden, although users can interact with
them directly through function arguments. Similarly, the image reading
and resizing module is fixed but will adapt to different Keras models by
checking their input shapes. The remaining seven modules are designed to
be overridable, enabling users to tailor the infrastructure to their
specific needs. These modules are discussed in detail on the software's
website.
